Mergeable Types

### 2017-02-28 paper/related.tex
edits

\name shares resemblance to conflict-free replicated data types
(CRDT)~\cite{crdt}. CRDTs define abstract data types such as counters, sets,
etc., with commutative operations such that the state of the data type always
converges. Unlike CRDTs, the operations on mergeable types in \name need not
commute and the reconciliation protocol is defined by user-defined merge
functions. Moreover, CRDTs are not composable and each CRDT tends to be built
from scratch over the network protocol. Compare to this, \name types are
composable; mergeable ropes are polymorphic over its mergeable content. \name
uses 3-way merges using the lowest common ancestor, which is critical for all
of our user-defined merges. However, CRDTs do not have the benefit of lowest
common ancestor for merges and are only presented with the two concurrent
versions. If a 3-way merge is desired, then the causal history has to be
explicitly encoded in the data type. As a result, constructing even simple data
types like counters are more complicated using CRDTs~\cite{crdt} compared to
their implementation in \name.

>>>>

\name shares resemblance to conflict-free replicated data types
(CRDT)~\cite{crdt}. CRDTs define abstract data types such as counters, sets,
etc., with commutative operations such that the state of the data type always
converges. Unlike CRDTs, the operations on mergeable types in \name need not
commute and the reconciliation protocol is defined by user-defined merge
functions. Moreover, CRDTs are not composable and each CRDT tends to be built
from scratch over the network protocol. Compare to this, \name types are
composable; mergeable ropes are polymorphic over its mergeable content. A
programmer working working with CRDTs is provided with a fixed set of mergeable
data types, with no recourse to constructing their own from the available
types.

\name uses 3-way merges using the lowest common ancestor, which is critical for
all of our user-defined merges. However, CRDTs do not have the benefit of
lowest common ancestor for merges and are only presented with the two
concurrent versions. If a 3-way merge is desired, then the causal history has
to be explicitly encoded in the data type. As a result, constructing even
simple data types like counters are more complicated using CRDTs~\cite{crdt}
compared to their implementation in \name. CRDTs also tend to be implemented
directly over the network protocols. Hence, low-level concerns such as
duplicate delivery, lost messages, message reordering are explicitly handled in
the data type definition. In \name, the programmer does not have to deal with
underlying network issues since \name is realized on Irmin, a high-level
branch-consistent distributed store.
